
# Reproducible Research - Assignment 1

set working directory and read file
```{r}
setwd("C:\\Coursera\\DataScience\\5_ReproducibleResearch\\Project1")
activ = read.csv("activity.csv")
```

calculate mean and median
```{r}
steps.mean = mean(activ$steps, na.rm = TRUE)
steps.median = median(activ$steps, na.rm = TRUE)
```

Create histogram for total steps per day
```{r}
library(ggplot2)
qplot(steps, data=activ, geom="histogram", main="Total steps per day", xlab="# steps", 
   ylab="Counts")
```   

Calculate interval means. First create intervals as factors
```{r}
facint = factor(activ$interval)
interval.ave <- aggregate(activ$steps, by = list(facint), FUN = mean, na.rm=TRUE)
names(interval.ave) = c("Interval", "Mean")
``` 

Time series plot
```{r}
plot(interval.ave$Interval, interval.ave$Mean, xlab= "Interval", ylab= "Mean", type = "l", col = "red") 
``` 

interval with maximum number of steps across all the days
```{r}
interval.ave[interval.ave$Mean == max(interval.ave$Mean),]
``` 

Total number of missing values
```{r}
sum(is.na(activ$steps))
``` 

Fill in NAs in new activfull dataframe with the mean for the interval 
```{r}
activfull = activ
for(i in seq_along(activfull$steps)) if(is.na(activfull$steps[i])) activfull$steps[i] <- interval.ave$Mean[interval.ave$Interval == activfull$interval[i]]
sum(is.na(activfull$steps))
``` 

histogram of new dataset
```{r}
qplot(steps, data=activfull, geom="histogram", main="Total steps per day (without NAs)", xlab="# steps", 
   ylab="Counts")
``` 

mean and median in new dataset
```{r}
stepsfull.mean = mean(activfull$steps, na.rm = TRUE)
stepsfull.median = median(activfull$steps, na.rm = TRUE)
``` 
create new dataset
```{r}
activfull2 = activfull
```

get day names
```{r}
activfull2$day = weekdays(as.Date(activfull2$date))
activfull2$day = as.factor(activfull2$day)
activfull2$daytype[activfull2$day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")] = c("W")
activfull2$daytype[activfull2$day %in% c("Saturday", "Sunday")] = "E" 
activfull2$daytype = as.factor(activfull2$daytype)
``` 

Calculate interval means. First create intervals as factors
```{r}
interval2.ave <- aggregate(activfull2$steps, by = list(facint, activfull2$daytype), FUN = mean, na.rm=TRUE)
names(interval2.ave) = c("Interval", "DayType", "Mean")
``` 

Create weekday and weekend factors
```{r}
interval2.ave$DayType = factor(interval2.ave$DayType, labels = c("Weekday", "Weekend"))
``` 

Plot weekend and weekdays 
```{r}
library(lattice)
xyplot(Mean ~ Interval | DayType, data = interval2.ave, type="l", layout = c(1, 2))
``` 






